---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Charlize Ezernack"
date: "01-31-2026"
output: github_document
---

### Load packages and data



```{r load-packages, message = FALSE}

library(tidyverse) 
devtools::install_github("rstudio-education/dsbox") 
library(dsbox) 

```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dennydata}
nrow(dennys)
ncol(dennys)
glimpse(dennys)
```

The dimensions are 1643 x 6, where each row indicates a location for a Denny's located indicated by the following variables: address, city, state, zip, longitude, and latitude.


### Exercise 2
```{r laquintadata}
nrow(laquinta)
ncol(laquinta)
glimpse(laquinta)
```

The dimensions are 909 x 6, where each row indicates a location for a La Quinta hotel indicated by the following variables: address, city, state, zip, longitude, and latitude.
Remove this text, and add your answer for Exercise 2 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

After looking at the La Quinta website there are locations outside the U.S in Asia, South America, the Middle East, Europe, Australia and the Pacific Rim, and Canada.

There does not seem to be any locations outside the U.S for Denny's.

In terms of looking at the data I believe I can try to filter for locations that are "NA" for state for La Quinta.

### Exercise 4

In terms of looking at the data I believe I can try to filter for locations that are "NA" for state for La Quinta.


### Exercise 5

```{r filterUSdennys}
dennys %>%
  filter(!(state %in% states$abbreviation))

dennys %>%
  mutate(country = "United States")
```

### Exercise 6

```{r mutateUSdennys}

dennys %>%
  mutate(country = "United States")
```

 
### Exercise 7

There are locations in Canada, Mexico, China, New Zealand, Georgia, Turkiye, United Arab Emirates, Colombia, and Ecudaor. 

### Exercise 8
```{r outsideUSlq}
laquinta %>%
  filter(!(state %in% states$abbreviation))

laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
     state %in% c("AG", "QR","CH", "PU", "NL", "VE", "SL" ) ~ "Mexico", state %in% c("ON", "BC") ~ "China"
  ))
```
### Exercise 9
```{r numUS}
laquinta <- laquinta %>%
  filter(country == "United States")

dennys %>% count(state)
laquinta %>% count (state)
```

California has the most Denny's locations at 403, and Texas trails slightly behind with 200. The states with low amounts of Denny's have 3 restaurants, the states are South Dakota, West Virginia, New Hampshire, Arkansas.  
Delaware has the lowest amount at 1 location. 

Texas has the most La Quinta locations by far at 237. Maine has the lowest amount of locations with only 1. 

I do not find this super surprising since I would assume popular tourist spots to have more locations of both La Quinta and Denny's.

### Exercise 10
```{r joindata}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

laquinta %>% 
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))


```
<!--put answers for #9, do area/n and then add it to a new variable to answer square miles question -->
### Exercise 11
```{r join}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
  
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
  
dn_lq <- bind_rows(dennys, laquinta)

ggplot(dn_lq, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point()
```
### Exercise 12
```{r plotNC}

data1 <- dn_lq %>% filter(state == "NC")

ggplot(data1, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) + labs(title = "Concentration of Denny's and La Quinta", subtitle = "In North Carolina") +
  geom_point(alpha= .5)
```

Yes the joke slightly holds here because every La Quinta is super close to a Denny's, shown by the blue and pink dots overlapping often. 

### Exercise 13
```{r plotTX}

data2 <- dn_lq %>% filter(state == "TX")

ggplot(data2, mapping = aes(
  x = longitude,
  y = latitude,
  color = establishment
)) + labs(title = "Concentration of Denny's and La Quinta", subtitle = "In Texas") +
  geom_point(alpha = .3)
```

Yes, the joke absolutely holds here. There are many clusters of pink and blue, depicting high concentration of both establishments in the same area. 