---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Charlize Ezernack"
date: "01-31-2026"
output: github_document
---

### Load packages and data



```{r load-packages, message = FALSE}

library(tidyverse) 
devtools::install_github("rstudio-education/dsbox") 
library(dsbox) 

```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dennydata}
nrow(dennys)
ncol(dennys)
glimpse(dennys)
```

The dimensions are 1643 x 6, where each row indicates a location for a Denny's located indicated by the following variables: address, city, state, zip, longitude, and latitude.


### Exercise 2
```{r laquintadata}
nrow(laquinta)
ncol(laquinta)
glimpse(laquinta)
```

The dimensions are 909 x 6, where each row indicates a location for a La Quinta hotel indicated by the following variables: address, city, state, zip, longitude, and latitude.
Remove this text, and add your answer for Exercise 2 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

After looking at the La Quinta website there are locations outside the U.S in Asia, South America, the Middle East, Europe, Australia and the Pacific Rim, and Canada.

There does not seem to be any locations outside the U.S for Denny's.

In terms of looking at the data I believe I can try to filter for locations that are "NA" for state for La Quinta.

### Exercise 4

In terms of looking at the data I believe I can try to filter for locations that are "NA" for state for La Quinta.


### Exercise 5

```{r filterUSdennys}
dennys %>%
  filter(!(state %in% states$abbreviation))

dennys %>%
  mutate(country = "United States")
```

### Exercise 6

```{r mutateUSdennys}

dennys %>%
  mutate(country = "United States")
```

 
### Exercise 7

There are locations in Canada, Mexico, China, New Zealand, Georgia, Turkiye, United Arab Emirates, Colombia, and Ecudaor. 

### Exercise 8
```{r outsideUSlq}
laquinta %>%
  filter(!(state %in% states$abbreviation))

laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
     state %in% c("AG", "QR","CH", "PU", "NL", "VE", "SL" ) ~ "Mexico", state %in% c("ON", "BC") ~ "China"
  ))
```
### Exercise 9
```{r filterlqUS}
lq <- lq %>%
  filter(country == "United States")
```
